### 2017-05-21 pds/Paper/pds_references.tex
updated references. added some ml description

%pds references

%Ethics/ Best Practices Papers & Articles
@misc{new2016,
  title = {{Living with the algorithms that run our lives} Kernel Description},
  howpublished = {\url{https://www.newscientist.com/article/mg22530073.500-living-with-the-algorithms-that-run-our-lives/
}},
  note = {Accessed: 2017-05-21}
}
@misc{algorithms2016,
  title = {{No one in control} Kernel Description},
  howpublished = {\url{http://thepoliticalgrill.yuku.com/topic/2328/No-one-in-control-The-algorithms-that-run-our-lives#.WSAL_TclHCI
}},
  note = {Accessed: 2017-05-21}
}
@misc{principles,
  title = {{Principles for accountable algorithms} Kernel Description},
  howpublished = {\url{%ttp://www.fatml.org/resources/principles-for-accountable-algorithms
}},
  note = {Accessed: 2017-05-21}
}

%DS Papers

%General ML Texts
@article{fayyad2001digital,
  title={The digital physics of data mining},
  author={Fayyad, Usama},
  journal={Communications of the ACM},
  volume={44},
  number={3},
  pages={62--62},
  year={2001},
  publisher={Association for Computing Machinery, Inc.}
}
@article{domingos2012few,
  title={A few useful things to know about machine learning},
  author={Domingos, Pedro},
  journal={Communications of the ACM},
  volume={55},
  number={10},
  pages={78--87},
  year={2012},
  publisher={ACM}
}
@inproceedings{feldman2015certifying,
  title={Certifying and removing disparate impact},
  author={Feldman, Michael and Friedler, Sorelle A and Moeller, John and Scheidegger, Carlos and Venkatasubramanian, Suresh},
  booktitle={Proceedings of the 21th ACM SIGKDD International Conference on Knowledge Discovery and Data Mining},
  pages={259--268},
  year={2015},
  organization={ACM}
}
@article{barocas2016big,
  title={Big data's disparate impact},
  author={Barocas, Solon and Selbst, Andrew D},
  year={2016}
}