### 2017-05-28 pds/Paper/pds_references.tex
advanced some in second part

%pds references

%Ethics/ Best Practices Papers & Articles
@misc{new2016,
  title = {{Living with the algorithms that run our lives} Kernel Description},
  howpublished = {\url{https://www.newscientist.com/article/mg22530073.500-living-with-the-algorithms-that-run-our-lives/
}},
  note = {Accessed: 2017-05-21}
}
@misc{algorithms2016,
  title = {{No one in control} Kernel Description},
  howpublished = {\url{http://thepoliticalgrill.yuku.com/topic/2328/No-one-in-control-The-algorithms-that-run-our-lives#.WSAL_TclHCI
}},
  note = {Accessed: 2017-05-21}
}
@misc{principles,
  title = {{Principles for accountable algorithms} Kernel Description},
  howpublished = {\url{%ttp://www.fatml.org/resources/principles-for-accountable-algorithms
}},
  note = {Accessed: 2017-05-21}
}

@misc{principles,
  title = {{Discrimination algorithms disparate impact} Kernel Description},
  howpublished = {\url{https://www.theatlantic.com/business/archive/2015/09/discrimination-algorithms-disparate-impact/403969/
}},
  note = {Accessed: 2017-05-22}
}
@PROCEEDINGS{proceedings,
  title = {Proc.\ o.t. First Workshop on Ethics in NLP
      (Valencia, Spain, April 2017)},
  booktitle = {Proc.\ o.t. First Workshop on Ethics in NLP
      (Valencia, Spain, April 2017)},
  volume = 1,
  series = {},
  publisher = {},
  address = {Valencia, Spain},
  year = {2017}
}

@inproceedings{leidner2017ethical,
  author =       {Jochen L. Leidner, Vassilis Plachouras},
  title =        {Ethical by Design: Ethics Best Practices for Natural Language Processing},
  crossref =     {proceedings},
  pages =        {9--18},
}

%DS Papers
@article{wagner2015s,
  title={It's a man's wikipedia? Assessing gender inequality in an online encyclopedia},
  author={Wagner, Claudia and Garcia, David and Jadidi, Mohsen and Strohmaier, Markus},
  journal={arXiv preprint arXiv:1501.06307},
  year={2015}
}
@article{harris1954distributional,
  title={Distributional structure},
  author={Harris, Zellig S},
  journal={Word},
  volume={10},
  number={2-3},
  pages={146--162},
  year={1954},
  publisher={Taylor \& Francis}
}
@article{boleda2016formal,
  title={Formal Distributional Semantics: Introduction to the Special Issue},
  author={Boleda, Gemma and Herbelot, Aur{\'e}lie},
  journal={Computational Linguistics},
  year={2016},
  publisher={MIT Press}
}
@article{chowdhury2003natural,
  title={Natural language processing},
  author={Chowdhury, Gobinda G},
  journal={Annual review of information science and technology},
  volume={37},
  number={1},
  pages={51--89},
  year={2003},
  publisher={Wiley Online Library}
}

%General ML Texts
@article{fayyad2001digital,
  title={The digital physics of data mining},
  author={Fayyad, Usama},
  journal={Communications of the ACM},
  volume={44},
  number={3},
  pages={62--62},
  year={2001},
  publisher={Association for Computing Machinery, Inc.}
}
@article{domingos2012few,
  title={A few useful things to know about machine learning},
  author={Domingos, Pedro},
  journal={Communications of the ACM},
  volume={55},
  number={10},
  pages={78--87},
  year={2012},
  publisher={ACM}
}
@inproceedings{feldman2015certifying,
  title={Certifying and removing disparate impact},
  author={Feldman, Michael and Friedler, Sorelle A and Moeller, John and Scheidegger, Carlos and Venkatasubramanian, Suresh},
  booktitle={Proceedings of the 21th ACM SIGKDD International Conference on Knowledge Discovery and Data Mining},
  pages={259--268},
  year={2015},
  organization={ACM}
}
@article{barocas2016big,
  title={Big data's disparate impact},
  author={Barocas, Solon and Selbst, Andrew D},
  year={2016}
}